### Task 2.  (Viktoria Kocherkevych)
#### 		In this task, we discuss a real-life process that is well modelled by a Poisson distribution. As you remember, a Poisson random variable describes occurrences of rare events, i.e., counts the number of successes in a large number of independent random experiments. One of the typical examples is the __radioactive decay__ process.
		
#### Consider a sample of radioactive element of mass $m$, which has a big _half-life period_ \(T\); it is vitally important to know the probability that during a one second period, the number of nuclei decays will not exceed some critical level \(k\). This probability can easily be estimated using the fact that, given the _activity_ ${\lambda}$ of the element (i.e., the probability that exactly one nucleus decays in one second) and the number $N$ of atoms in the sample, the random number of decays within a second is well modelled by Poisson distribution with parameter $\mu:=N\lambda$. Next, for the sample of mass $m$, the number of atoms is $N = \frac{m}{M} N_A$, where $N_A = 6 \times 10^{23}$ is the Avogadro constant, and $M$ is the molar (atomic) mass of the element. The activity of the element, $\lambda$, is $\log(2)/T$, where $T$ is measured in seconds. 

#### Assume that a medical laboratory receives $n$ samples of radioactive element ${{}^{137}}\mathtt{Cs}$ (used in radiotherapy) with half-life period $T = 30.1$ years and mass \(m = \mathtt{team\, id \,number} \times 10^{-6}\) g each. Denote by $X_1,X_2,\dots,X_n$ the __i.i.d.  r.v.__'s counting the number of decays in sample $i$ in one second. 

1.  Specify the parameter of the Poisson distribution of \(X_i\) (you'll need the atomic mass of _Cesium-137_)  
2.  Show that the distribution of the sample means of \(X_1,\dots,X_n\) gets very close to a normal one as $n$ becomes large and identify that normal distribution. To this end,
    +  simulate the realization $x_1,x_2,\dots,x_n$ of the $X_i$ and calculate the sample mean $s=\overline{\mathbf{x}}$;
    +  repeat this $K$ times to get the sample $\mathbf{s}=(s_1,\dots,s_K)$ of means and form the empirical cumulative distribution function \(\hat  F_{\mathbf{s}}\) of $\mathbf{s}$;
    +  identify $\mu$ and $\sigma^2$ such that the \textbf{c.d.f.} \(F\) of $\mathscr{N}(\mu,\sigma^2)$ is close to the \textbf{e.c.d.f.} \(\hat F_{\mathbf{s}}\) and plot both __c.d.f.__'s on one graph to visualize their proximity (use the proper scales!);
    +  calculate the maximal difference between the two \textbf{c.d.f.}'s;
    +  consider cases $n = 5$, $n = 10$, $n=50$ and comment on the results.   
3.  Calculate the largest possible value of $n$, for which the total number of decays in one second is less than $8 \times 10^8$ with probability at least $0.95$. To this end,  
    +  obtain the theoretical bound on \(n\) using Markov inequality, Chernoff bound and Central Limit Theorem, and compare the results;  
    +  simulate the realization $x_1,x_2,\dots,x_n$ of the $X_i$ and calculate the sum $s=x_1 + \cdots +x_n$;
    +  repeat this $K$ times to get the sample $\mathbf{s}=(s_1,\dots,s_K)$ of sums;
    +  calculate the number of elements of the sample which are less than critical value ($8 \times 10^8$) and calculate the empirical probability; comment whether it is close to the desired level \(0.95\)

```{r}

m <- id * (10^(-6))  #grams
Na <- 6 * (10^23)
M <- 139
T_y <- 30.1 #years
T_s <- T_y * 365 * 24 * 3600

lambda <- log(2) / T_s  
N <- m*Na/M

puasson <- N * lambda
puasson
K <- 1e3
n <- 5

sample_means <- colMeans(matrix(rpois(n*K, lambda = puasson), nrow=n))

s_mean <- mean(sample_means)

means <- replicate(K, mean(colMeans(matrix(rpois(n*K, lambda = puasson), nrow=n))))



```


#### Next, calculate the parameters of the standard normal approximation

```{r}

mu <- mean(means)
sigma <- sd(means)
mu
sigma
```

#### We can now plot ecdf and cdf

```{r}

xlims <- c(mu-3*sigma,mu+3*sigma)
Fs <- ecdf(means)
plot(Fs,
  ylim = c(0,1),
  col = "blue",
  lwd = 2,
  main = "Comparison of ecdf and cdf")
curve(pnorm(x, mean = mu, sd = sigma), col="red", lwd=2, add=TRUE)
x <- seq(min(means), max(means), by=0.01)
diff <- max(abs(Fs(x) - pnorm(x, mean = mu, sd= sigma)))
diff
```
##### Theoretical bound on n:

###### **Markov inequality**:
Let S be the sum of $x_1,x_2,\dots,x_n$. Then $ P(S \le 8 * 10^8) = 1 - P(S \ge8*10^8) \le \frac{1}{8*10^8} E(S) \ge 0.95$ 

$ \frac{1}{8*10^8} E(S) \le 0.05$

$ E(S) = \mu*n $

$ n \le \frac{4*10^7}{\mu} $

$ n \le \frac{4*10^7}{59888359}$

$ n \le 0.667 $

$ n = 0 $


###### **Chernoff bound**:

$ P(S \le 8*10^8) \le  $


###### **Central Limit Theorem**:

$ P(\frac{S_{n}-n\mu}{\sigma\sqrt{n}} \le \frac{8*10^8-n\mu}{\sigma\sqrt{n}}) \rightarrow \Phi(t) $

$ \Phi(t) = \Phi(\frac{8*10^8-n\mu}{\sigma\sqrt{n}})\ge 0.95 $

$ \Phi(t) $ has to be equal to $ 1.65 $ (due to the Standart normal table)

(https://www.sjsu.edu/faculty/gerstman/EpiInfo/z-table.htm)

After all the calculations we have:

$ n \approx 13.157 $

$ n = 13 $

Summary: Markov inequality is the most unreliable of all the theorems.

```{r}

n<-13
sample_sums <- colSums(matrix(rpois(13*K, lambda = puasson), nrow=n))

good_sums <- length(which(samples>=8*(10^8)))

good_sums


```

